---
title: "Nov 17, 2019 Matrix Operation Basic"
author: "Jikhan Jeong"
date: "11/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matrix Operation in R
* Ref: http://www.philender.com/courses/multivariate/notes/matr.html
* ctrl + Enter         : run this line
* ctrl + shift + Enter : run this chuck
* Ctrl + Shift + K     : Knit
* Ctrl + Alt   + I     : Insert a chuck

# Making matrix by using matrix
```{r}
A <- matrix(c(seq(1:6)), 2,3)
A
```

```{r}
is.matrix(A)
```
```{r}
is.vector(A)
```

```{r}
B <- matrix(c(seq(1:6)), 2,3)
C <- A+B
C
```
# Transpose t(matrix)
```{r}
t(C)
```
# Matrix Multiplication
```{r}
A # 2 x 3
t(C) # traspose 3 X 2  
A%*%t(C) # 2 x x
```
# Unit Vector
```{r}
(U <- matrix(1,2,2))
```
# Zero Vector
```{r}
(Z <- matrix(0,2,2))
```
# Diagonal component
```{r}
C
diag(C)
diag(diag(C))
```
# Identity Matrix
```{r}
c(1,1,1)
(I <-diag(c(1,1,1)))
```
# Inverse of a Matrix
```{r}
(A <- matrix(c(4,4,-2,2,6,2,2,8,4),3,3))
(inv.A <- solve(A))
A%*%inv.A 
inv.A %*%A 
```
# Inverse & Determinant of a Matrix (=det(A))
```{r}
A
solve(A)
det(A)
```
# Rank of a Matrix
* Full rank condition
```{r}
A <- matrix(c(2,3,-2,1,2,2,4,7,0),3,3)
matA <- qr(A)
matA$rank
```
* Under justification
* column 3 is 2 times column 1
```{r}
A <- matrix(c(2,3,-2,1,2,2,4,6,-4),3,3)
A
matA <- qr(A)
matA$rank
```
# Number of Rows & Columns
```{r}
X <- matrix(c(3,2,4,3,2,-2,6,1),4,2)
X
dim(X)
(r <-nrow(X))
(c<-ncol(X))

```
# Computing Column & Row Sums
```{r}
(A <- matrix(c(2,3,-2,1,2,2),3,2))
(c<- colSums(A))
(r<- rowSums(A))
(a<- sum(A))
```
# Computing Column & Row Means
```{r}
A
(cm <- colMeans(A))
(rm <- rowMeans(A))
(m  <- mean(A))

```
# Horizontal Concatenation: A+B **cbind**
```{r}
A
(B <- matrix(c(1,3,2,1,4,2),3,2))
(C <-cbind(A,B))
```
# Vertical Concatenation (Appending): A + B(next row) **rbind**  
```{r}
(c<-rbind(A,B))
```

